Concurrent, Parallel Garbage Collection in Linear Time

### 2016-06-01 finaletd/mainintroduction.tex
intro

Beyong solving the garbage collection in concurrent programming environment, the main objectives of this thesis are to design garbage collectors for shared and distributed memory system that satisfy the following properties
\begin{enumerate}
	\item Concurrent garbage collectors (Less pause time)
	\item Multi-collector garbage collection
	\item No global synchornization (High Throughput)
	\item Locality based garbage collection
	\item Scalable
	\item Promptness
	\item Safety
	\item Completeness
\end{enumerate}

\paragraph{Concurrent:}
	Concurrent garbage collectors have negligible to zero pause time. As these type of collectors process heap along side application, application does not get locked and waiting for the garbage collector to finish processing.
\paragraph{Multi-collector Garbage :}
	Multi-collector garbage collections are multiple independent garbage collectors that work independently on the heap to identify garabge. These collectors utilize multi-processors systems when there is lot of compute resources being under utilized.
\paragraph{Global Synchronization:}
	When multi-collector garbage collectors are used, conventional solutions require global barrier among all the collectors to communicate and share the information to identify garbage objects and delete them. With no global synchronization, throughput of garbage collectors will be high.
\paragraph{Locality based garbage collection:}
	Detecting garbage is a global predicate. Traditionally, popular methods of garbage collection algorithm involves computing garbage as global predicate. This requires scanning entire heap for better evaluating the predicate. If garbage object can be detected locally based on the a small set of object, then the garbage detection process can be parallelized.
\paragraph{Scalable:}
	With increase in the size of memory, conventional global garbage collection techniques will not scale with synchronization and scanning the entire heap(s). A garbage collection algorithm must be scalable to meet the future demands. Locality based garbage collection with no globally synchronized multi-collector mechanism is the way for scalability.
\paragraph{Promptness:}
	With global garbage collection, the cost of scanning entire heap often is very expensive. It comes with the cost that the garbage will not be collected until the scanning process starts. We require garbage collector that are prompt in collecting garbage as the heap(s) scanning might be expensive in huge distributed systems.
\paragraph{Safety:}
	Safety property is the most crucial part of the problem. Every garbage collector must delete only garbage object. While most garbage collector in literature satisfy this property, there are some that cannot satisfy this property. This thesis requires garbage collector to be safety in the multi-collector environment which is challenging given that there is no global synchronization.
\paragraph{Completeness:}
	Completeness helps garbage collector to claim that it collects all garbage objects in the heap. Reference counting garbage collector is well known for incomplete garbage colletion due to their inability to collect cyclic garbage. This property is challenging as the locality based collectors usually use reference counting. Garbage collector with localized decision making guarantee completeness is necessary to make the collector highly adaptable to enhance the performance of the applications.
\section{Contributions}
\label{intro:contr}
\section{Dissertation Organization}
\label{intro:do}

>>>>

Beyond solving the garbage collection in concurrent programming environment, the main objectives of this thesis are to design garbage collectors for shared and distributed memory system that satisfy the following properties :
\begin{enumerate}
	\item Concurrent garbage collectors (Less pause time)
	\item Multi-collector garbage collection
	\item No global synchronization (High Throughput)
	\item Locality based garbage collection
	\item Scalable
	\item Promptness
	\item Safety
	\item Completeness
\end{enumerate}

There are no garbage collection algorithms available for shared and distributed memory systems that satisfy all the above mentioned properties.  This thesis focuses on designing novel garbage collector with significant improvement. Apart from garbage collection, the ideas presented in the thesis will be of use to solve other theoretical problems in distributed computing like dynamic reachability, breaking cycle, detecting cycles in graph, data aggregation in the sensor network, broken network detection in sensor networks.

\paragraph{Concurrent:}
	Concurrent garbage collectors work simultaneously with application. Application will not experience any pause during execution to scan the memory. They
	have negligible to zero pause time. There are handful of collectors that satisfy this property. 
\paragraph{Multi-collector Garbage :}
	Multi-collector garbage collectors are multiple independent garbage collector thread / process that work independently on the heap to identify garabge. These collectors utilize multiple processors when there is lot of compute resources being under utilized. These type of collectors are essential in distributed systems and shared memory systems. In shared memory systems, when multiple processors are under utilized, this pro
\paragraph{Global Synchronization:}
	When multi-collector garbage collectors are used, conventional solutions require global barrier among all the collectors to communicate and share the information to identify garbage objects and delete them. With no global synchronization, throughput of garbage collectors will be high.
\paragraph{Locality based garbage collection:}
	Detecting garbage is a global predicate. Traditionally, popular methods of garbage collection algorithm involves computing garbage as global predicate. This requires scanning entire heap for better evaluating the predicate. If garbage object can be detected locally based on the a small set of object, then the garbage detection process can be parallelized.
\paragraph{Scalable:}
	With increase in the size of memory, conventional global garbage collection techniques will not scale with synchronization and scanning the entire heap(s). A garbage collection algorithm must be scalable to meet the future demands. Locality based garbage collection with no globally synchronized multi-collector mechanism is the way for scalability.
\paragraph{Promptness:}
	With global garbage collection, the cost of scanning entire heap often is very expensive. It comes with the cost that the garbage will not be collected until the scanning process starts. We require garbage collector that are prompt in collecting garbage as the heap(s) scanning might be expensive in huge distributed systems.
\paragraph{Safety:}
	Safety property is the most crucial part of the problem. Every garbage collector must delete only garbage object. While most garbage collector in literature satisfy this property, there are some that cannot satisfy this property. This thesis requires garbage collector to be safety in the multi-collector environment which is challenging given that there is no global synchronization.
\paragraph{Completeness:}
	Completeness helps garbage collector to claim that it collects all garbage objects in the heap. Reference counting garbage collector is well known for incomplete garbage colletion due to their inability to collect cyclic garbage. This property is challenging as the locality based collectors usually use reference counting. Garbage collector with localized decision making guarantee completeness is necessary to make the collector highly adaptable to enhance the performance of the applications.
\section{Contributions}
\label{intro:contr}
\section{Dissertation Organization}
\label{intro:do}
